---
title: "Interpolation"
author: "Aron Lesser"
date: "10/30/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(stringr)
```


# Adding Data

I read the downloaded census tract file with median income. Here I select the geoid and med.income only. I then mutate the geo_id into geoid10 which is the name on the other dataset, transform it into numeric, and isolate the last 11 numbers . 

```{r warning=FALSE}
tracts_info <- st_read("income_boston_tracts.csv", quiet = TRUE) %>%
  dplyr::select(GEO_ID, S1901_C01_012E) %>%
  mutate(GEOID10 = as.numeric (str_sub(GEO_ID,-11,-1))) %>%
  dplyr::select(GEOID10, S1901_C01_012E) 
  
  

```

Here I make geoid 10 into a numeric so that they can be joined
```{r message=FALSE}
census_tracts <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/4a8eb4fb3be44ed5a1eec28551b9f3b2_0.geojson?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D", quiet = TRUE) %>%
  mutate (GEOID10 = as.numeric(GEOID10))

```

I create a new dataset with the left join 
```{r warning=FALSE}
income <- left_join(census_tracts, tracts_info, by = "GEOID10", copy = FALSE) %>%
  mutate(S1901_C01_012E = case_when(S1901_C01_012E == "-" ~ NA_real_,
                                            S1901_C01_012E != "-" ~ as.numeric(S1901_C01_012E))) %>%
  mutate(S1901_C01_012E = case_when(is.na(S1901_C01_012E) ~ round(mean(S1901_C01_012E, na.rm = TRUE),1),
                                          !is.na(S1901_C01_012E) ~ S1901_C01_012E)) 
  

income = income[-48,]
```


# Mapping median income in boston with a choropleth

```{r warning=FALSE}

income$describe <- 
  paste("Tract", income$OBJECTID, "<br>", "Median Income", "$",prettyNum(income$S1901_C01_012E, digits = 4))%>% 
  lapply(htmltools::HTML)

bins <- seq(min(income$S1901_C01_012E),
            max(income$S1901_C01_012E), by = 20000)
pal <- colorNumeric("viridis", 
                    domain = income$S1901_C01_012E,
                    na.color = "#00000000")

leaflet(income) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = income$describe,
              fillColor = ~pal(S1901_C01_012E),
              weight = 1, color = "black") %>% 
  addLegend(pal = pal, 
            values = ~S1901_C01_012E,
            bins = 10,
            opacity = 0.7, title = "Median Income",
            position = "topright")
```

# Mapping median income as points
```{r warning=FALSE}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

income_points <- st_centroid(
  st_transform(income, crs = MA_state_plane)) %>%
  st_transform(WGS84)

leaflet(income_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addCircles(highlightOptions = highlightOptions(fillOpacity = 1),
              label = income$describe,
              fillColor = ~pal(S1901_C01_012E),
             stroke = FALSE,
              radius = 50,
             fillOpacity = 1) %>% 
  addLegend(pal = pal, 
            values = ~S1901_C01_012E,
            bins = 10,
            opacity = 0.7, title = "Median Income",
            position = "topright")
```


Convert to spatial

```{r warning=FALSE}
income_points_sp <- income_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

income_poly_sp <- income %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()
```

```{r warning=FALSE}
boston_raster <- raster(income_poly_sp, res=10)
```

```{r warning=FALSE}
gs <- gstat(formula=S1901_C01_012E~1, locations=income_points_sp)
idw_interp <- interpolate(boston_raster, gs)

idw_interp_clip <- mask(idw_interp, income_poly_sp)
```

# Mappping interpolation median income

```{r warning=FALSE}
leaflet(income_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>% 
  addLegend(pal = pal, 
            values = ~S1901_C01_012E,
            bins = 10,
            opacity = 0.7, title = "Median Income",
            position = "topright")
```

#Discussion

While I see value in usign an interpolation for certain circumstances, it is not necessarily a better tool than a simple choropleth map. It would be interesting to do the same exercise with neighborhood - rather than census tract - income data. If I only had a few data points for income, then interpolation would be a more appropriate method. 

