---
title: "Vis Assignment 3"
author: "Aron Lesser"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries, include=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```

```{r}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

openspace <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2868d370c55d4d458d4ae2224ef8cddd_7.kml", quiet =TRUE)

public_schools <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

n_schools <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/0046426a3e4340a6b025ad52b41be70a_1.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

```


Transforming the data

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"


nhoods <- nhoods %>%
  st_transform(MA_state_plane)

openspace <- openspace %>%
  st_transform(MA_state_plane)

public_schools <- public_schools %>%
  st_transform(MA_state_plane)

n_schools <- n_schools %>%
  st_transform(MA_state_plane)

```




```{r}
ggplot(openspace) +
  geom_sf(fill = "lightgreen", color = NA) +
  geom_sf(data = public_schools, color = "blue", size = 0.01) +
  geom_sf(data = n_schools, color = "black", size = 0.01) +
  geom_sf(data= nhoods, fill = NA, color = "gray") +
  theme_map() +
  annotation_scale()
```


Creating a buffer around public and private schools

```{r}
open_buffer <- st_buffer(openspace, dist = 100) %>%
  st_union()

ggplot(open_buffer) +
  geom_sf() +
  theme_map()
```

Subsetting parks and public schools

```{r}
public_ospace <- public_schools[open_buffer,]
  
ggplot(open_buffer) +
  geom_sf() +
  geom_sf(data = public_ospace, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()
```

Joining datasets

```{r}
public_schools <- public_schools %>%
  st_join(public_ospace) %>%
  mutate(by_openspace = !is.na(Name.y))
```


Calculating how many public schools are within 100 meters of an open space


```{r}
n_openspace_public <- sum(public_schools$by_openspace)

n_openspace_public
```

Percent of all public schools

```{r}
n_public <- length(public_schools$by_openspace)

pct_openspace_public <- n_openspace_public / n_public

pct_openspace_public
```

Nearly 60% of public schools are located within 100 meters of a park.




Mapping it 

```{r}
left_side  <- st_bbox(public_schools)$xmin
top_side <- st_bbox(public_schools)$ymax

ggplot(openspace) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = public_schools, size = 0.01,
          aes(color = by_openspace)) +
  scale_color_manual(values = c("lightgreen", "darkgreen"),
          name = "Boston Public Schools\nby distance to an open space", 
          labels = c("No open space within 100 m",
                     "Open space within 100 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_public, big.mark = ","),
                         " public schools in Boston\n", 
                         prettyNum(n_openspace_public, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_openspace_public, digits = 0),
                         "%) are within 100\nmeters of an open space",
                         sep = ""),
           hjust = 0, vjust = 0.75, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.7, 0.2))
```


# Chloropleth number of public schools in each neighborhood

```{r}
nhoods <- nhoods %>%
  mutate(num_public_schools = lengths(st_covers(nhoods, public_schools)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_public_schools)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of public schools",
                       breaks = breaks <- seq(0, 30, by = 6),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "public schools")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```





# Chloropleth number of private schools in each neighborhood

```{r}
nhoods <- nhoods %>%
  mutate(num_n_schools = lengths(st_covers(nhoods, n_schools)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_n_schools)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of private schools",
                       breaks = breaks <- seq(0, 30, by = 6),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "private schools")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```


















```{r fig.width=8, fig.height=8}
ggplot(neighborhoods) +
  geom_sf(aes(fill=Name)) +
geom_sf(data = schools) +
theme_map()
theme(legend.position = c(0.7, 0))

```


Count schools in each neighborhood

```{r}
test <- st_covers(neighborhoods, schools)

lengths(test)
```

```{r}
neighborhoods <- neighborhoods >%>
  
```

