---
title: "Vis Assignment 3"
author: "Aron Lesser"
date: "9/25/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries, include=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
```


# Loading Data

I loaded 4 sets of data for Boston: 

1. Boston neighborhood boundaries
2. Open space in Boston boundaries
3. Public Schools 
4. Private Schools

```{r}
nhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

openspace <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2868d370c55d4d458d4ae2224ef8cddd_7.kml", quiet =TRUE)

public_schools <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

n_schools <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/0046426a3e4340a6b025ad52b41be70a_1.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

```


# Transforming the data

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"


nhoods <- nhoods %>%
  st_transform(MA_state_plane)

openspace <- openspace %>%
  st_transform(MA_state_plane)

public_schools <- public_schools %>%
  st_transform(MA_state_plane)

n_schools <- n_schools %>%
  st_transform(MA_state_plane)

```


# Plotting everything on a map 

```{r}
ggplot(openspace) +
  geom_sf(fill = "lightgreen", color = NA) +
  geom_sf(data = public_schools, color = "blue", size = 0.01) +
  geom_sf(data = n_schools, color = "black", size = 0.01) +
  geom_sf(data= nhoods, fill = NA, color = "gray") +
  theme_map() +
  annotation_scale()
```


# Calculating Proximity of Public and Private Schools to Parks


## Public Schools

I began Creating a 100 meter buffer around open spaces. This will allow me to assess how many schools are in close proximity to open spaces. 

```{r}
open_buffer <- st_buffer(openspace, dist = 100) %>%
  st_union()

ggplot(open_buffer) +
  geom_sf() +
  theme_map()
```

I then subset the open spaces and public schools

```{r}
public_ospace <- public_schools[open_buffer,]
  
ggplot(open_buffer) +
  geom_sf() +
  geom_sf(data = public_ospace, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()
```

Next, I join those datasets

```{r}
public_schools <- public_schools %>%
  st_join(public_ospace) %>%
  mutate(by_openspace = !is.na(Name.y))
```


Here, I Calculate how many public schools are within 100 meters of an open space


```{r}
n_openspace_public <- sum(public_schools$by_openspace)

n_openspace_public
```

I then calculate what percent of all public schools those within 100 meters of open spaces represent.

```{r}
n_public <- length(public_schools$by_openspace)

pct_openspace_public <- n_openspace_public / n_public

pct_openspace_public
```

Nearly 60% of public schools are located within 100 meters of an open space



### Mapping Public Schools and Open Spaces 

```{r}
left_side  <- st_bbox(public_schools)$xmin
top_side <- st_bbox(public_schools)$ymax

ggplot(openspace) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = public_schools, size = 0.01,
          aes(color = by_openspace)) +
  scale_color_manual(values = c("lightgreen", "darkgreen"),
          name = "Boston Public Schools\nby distance to an open space", 
          labels = c("No open space within 100 m",
                     "Open space within 100 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_public, big.mark = ","),
                         " public schools in Boston\n", 
                         prettyNum(n_openspace_public, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_openspace_public, digits = 0),
                         "%) are within 100\nmeters of an open space",
                         sep = ""),
           hjust = 0, vjust = 0.75, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.7, 0.2))
```

## Private Schools and open spaces


I subset open spaces and private schools

```{r}
private_ospace <- n_schools[open_buffer,]
  
ggplot(open_buffer) +
  geom_sf() +
  geom_sf(data = private_ospace, 
          color = "darkgreen", 
          size = 0.01) +
  theme_map()
```

Next, I join the datasets for private schools

```{r}
n_schools <- n_schools %>%
  st_join(private_ospace) %>%
  mutate(by_openspace = !is.na(Name.y))
```


I then calculate how many private schools are within 100 meters of an open space


```{r}
n_openspace_private <- sum(n_schools$by_openspace)

n_openspace_private
```

I calculate the percent of all private schools that those within 100 meters of an open space represent.

```{r}
n_private <- length(n_schools$by_openspace)

pct_openspace_private <- n_openspace_private / n_private

pct_openspace_private
```

67% of private schools are located within 100 meters of a park.


### Mapping Private Schools and Open Spaces

```{r}
left_side  <- st_bbox(n_schools)$xmin
top_side <- st_bbox(n_schools)$ymax

ggplot(openspace) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = n_schools, size = 0.01,
          aes(color = by_openspace)) +
  scale_color_manual(values = c("green", "darkgreen"),
          name = "Boston Private Schools\nby distance to an open space", 
          labels = c("No open space within 100 m",
                     "Open space within 100 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side, 
           y = top_side, 
           label = paste("Of the ", 
                         prettyNum(n_public, big.mark = ","),
                         " private schools in Boston\n", 
                         prettyNum(n_openspace_private, big.mark = ","),
                         " (", 
                         prettyNum(100*pct_openspace_private, digits = 0),
                         "%) are within 100\nmeters of an open space",
                         sep = ""),
           hjust = 0, vjust = 0.2, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.7, 0.2))
```


The results from the analyses demonstrate that in Boston, a larger percentage of private schools are within close proximity to an open space (67%) than public schools (60%). 


# Choropleth map of the number of public schools in each neighborhood

```{r}
nhoods <- nhoods %>%
  mutate(num_public_schools = lengths(st_covers(nhoods, public_schools)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_public_schools)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of public schools",
                       breaks = breaks <- seq(0, 30, by = 6),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "public schools")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.7, 0.1))
```


# Choropleth map of the number of private schools in each neighborhood

```{r}
nhoods <- nhoods %>%
  mutate(num_n_schools = lengths(st_covers(nhoods, n_schools)))

ggplot(nhoods) +
  geom_sf(color = NA, 
          aes(fill = num_n_schools)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of private schools",
                       breaks = breaks <- seq(0, 30, by = 6),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "private schools")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.7, 0.1))
```

The largest neighborhood in the city, Dorchester, has the most private and public schools. There are notable differences in some northern neighborhoods, which seem to have a greater number of public schools. I also notice differences in some central neighborhoods, which have disproportionately more public schools than private schools.

# Average public school distance from open space

Here I calculate the average distance of public schools from open spaces. 

```{r results='hide'}
public_schools <- public_schools %>%
  mutate(open_dist = st_nn(public_schools, openspace, 
                           returnDist = TRUE)$dist) %>%
  mutate(open_dist = as.numeric(open_dist))

avg_open_dist <- mean(public_schools$open_dist)

avg_open_dist
```

I hide the results, which had a long string of lines and percentages. The result is 113.8929.


```{r fig.width=8, fig.height=8}
right_side <- st_bbox(public_schools)$xmax
left_side  <- st_bbox(public_schools)$xmin
top_side <- st_bbox(public_schools)$ymax
bottom_side <- st_bbox(public_schools)$ymin


ggplot(openspace) +
  geom_sf(fill = "lightgreen", color = NA) +
  geom_sf(data = public_schools, size = 0.1,
          aes(color = open_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "Boston Public Schools \nby distance to an open space (meters) ") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 300, 
           y = top_side - 550, 
           label = paste("On average, a Boston public school \nis ", 
                         prettyNum(avg_open_dist, digits = 3),
                         " meters from an open space",
                         sep = ""),
           hjust = 0, vjust = .6, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.63, 0.03))
```




# Private school distance from open space

I calculate the average distance of private schools from open spaces. 

```{r results='hide'}
n_schools <- n_schools %>%
  mutate(open_dist = st_nn(n_schools, openspace, 
                           returnDist = TRUE)$dist) %>%
  mutate(open_dist = as.numeric(open_dist))

avg_open_dist_private <- mean(n_schools$open_dist)

avg_open_dist_private

```
I hide the results, which had a long string of lines and percentages. The result is 93.04991.


Mapping Private School Distance from Open Space


```{r fig.width=8, fig.height=8}
right_side <- st_bbox(n_schools)$xmax
left_side  <- st_bbox(n_schools)$xmin
top_side <- st_bbox(n_schools)$ymax
bottom_side <- st_bbox(n_schools)$ymin


ggplot(openspace) +
  geom_sf(fill = "lightgreen", color = NA) +
  geom_sf(data = n_schools, size = 0.1,
          aes(color = open_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "Boston Private Schools \nby distance to an open space (meters) ") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 300, 
           y = top_side - 550, 
           label = paste("On average, a Boston Private school \nis ", 
                         prettyNum(avg_open_dist_private, digits = 3),
                         " meters from an open space",
                         sep = ""),
           hjust = 0, vjust = .6, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"),
        legend.position = c(0.65, 0.03))
```

In Boston, these findings suggest that private schools are nearly 10% closer to open spaces than public schools. 


# Calculating the percentage of land that open spaces represent in Dorchester and South Boston

## Area of neighborhoods 

Here, I add the areas in square kilometers to my datasets
```{r}


openspace <- openspace %>%
   mutate(openarea = set_units(st_area(openspace), km^2)) 



nhoods <- nhoods %>%
  mutate(area = set_units(st_area(nhoods), km^2))




```

## Finding the amount of land that is open space in Dorchester

I calculate the area of open spaces within the subset of Dorchetser. 


```{r}
Dorch <- nhoods[22,]

subset_parks_dorchester <- openspace[Dorch,]

sum(subset_parks_dorchester$openarea)

```

There are 2.86 km2 of open space in Dorchester, whose land area is 18.87129 km2. Therefore, 15.2% of Dorchester is occupied by open spaces. 


# Finding the amount of land that is open space in South Boston Waterfront

```{r}
SBW <- nhoods[23,] 

subset_parks_SBW <- openspace[SBW,]

sum(subset_parks_SBW$openarea)


```

#Finding the amount of land that is open space in South Boston 

I combine the two because I consider both South Boston and South Boston Waterfront to be part of the same neighborhood. 

```{r}
SB <- nhoods[24,] 

subset_parks_SB <- openspace[SB,]

sum(subset_parks_SB$openarea)


```

The area of open space in SBW is .083 km2
The area of open space in SB is .82 km 2
Therefore, the amount of open space in South Boston is .903 km2

The area of SBW is 2.52km2
The area of SB is 5.83 km2
Therefore, the area of South Boston is 8.35km2

Thus, 11% of South Boston is occupied by open spaces.


As a note, I imagine there is a more streamlined way to calculate the neighborhood areas that are occupied by open spaces. I would like to find a more efficient way to do so. 
